### 2022-06-24 manuscript/appendix-D-interaction-plots.Rmd

# Appendix D: Further plots of interactions {#appendix-D-interaction-plots}

```{r}

# This appendix inherits the libraries, the working directory and the
# data sets loaded in manuscript.Rmd, when the latter is knitted.

```

This appendix presents plots of interactions that did not fit in the main text.

## Study 1: Semantic priming

```{r semanticpriming-interactions-with-gender, fig.cap = 'Interactions with gender in the semantic priming study.', out.height = '70%'}

# Run plot through source() rather than directly in this R Markdown document 
# to preserve the italicised text.

source('semanticpriming/frequentist_analysis/semanticpriming-interactions-with-gender.R', 
       local = TRUE)

include_graphics(
  paste0(
    getwd(),  # Circumvent illegal characters in file path
    '/semanticpriming/frequentist_analysis/plots/semanticpriming-interactions-with-gender.pdf'
  ))

```

... TBC

## Study 2: Semantic decision

...

## Study 3: Lexical decision

...