### 2022-06-29 slides/slides.Rmd

---
title: "<img src='img/LU logo.jpg' alt='Lancaster University logo' width='130px' style='float:right; margin-top:-5%;'/><br>Language and vision in conceptual processing: Multilevel analysis and statistical power"
subtitle: 'Pablo Bernabeu'
author: 'Supervisors: &nbsp;Dermot Lynott and Louise Connell<br><br>Materials: http://doi.org/10.17605/OSF.IO/UERYQ'
institute: 'Lancaster University'
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      titleSlideClass: [middle, center]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

---

class: middle, left
layout: true

---

<!-- Load Bootstrap styles -->
<link href="https://cdn.jsdelivr.net/npm/bootstrap@5.0.2/dist/css/bootstrap.min.css" rel="stylesheet" integrity="sha384-EVSTQN3/azprG1Anm3QDgpJLIm9Nao0Yz1ztcQTwFspd3yD65VohhpuuCOmLASjC" crossorigin="anonymous">

<!-- Define CSS styles -->
<style>

.title-slide {
  background-color: #FAFAFC;
}
.title-slide h1,
.title-slide h2,
.title-slide h3,
.title-slide h4 {
  color: black;
  text-shadow: none;
}

/* Disable default striping in tables */
.remark-slide thead, 
.remark-slide tfoot, 
.remark-slide tr:nth-child(even) {
  background: none;
}

</style>

```{r setup, include = FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE, error = FALSE, 
                      echo = FALSE, knitr.duplicate.label = 'allow', 
                      fig.align = 'center')

library(dplyr)        # data wrangling
library(formattable)  # number formatting
library(kableExtra)   # table formatting

# Move working directory to the project root
knitr::opts_knit$set(root.dir = '../')

```

```{r xaringanExtra, include = FALSE}

xaringanExtra::use_xaringan_extra(c('tile_view', 'share_again', 'scribble',
                                    'logo', 'search', 'fit_screen'))

```

```{r load-functions-data-models, include = FALSE}

# Read in custom functions. Working directory has to be changed, but at the end 
# of this chunk, it is automatically reset to the manuscript.
setwd('R_functions')
sapply(list.files(), source, echo = FALSE)
setwd('../')

# Load main data sets and models. These objects are loaded directly, rather than 
# being run on the go, to keep the knitting of the manuscript reasonably fast. 
# Crucially, however, all the objects can be reproduced from the appropriate R 
# scripts in the current project.

# Study 1: Semantic priming
# Code for data set below in 'semanticpriming/data/data_preparation' folder
semanticpriming = read.csv('semanticpriming/data/final_dataset/semanticpriming.csv')
# Code for models below in 'analysis' folder
semanticpriming_lmerTest =   # Primary model object
  readRDS('semanticpriming/frequentist_analysis/results/semanticpriming_lmerTest.rds')
KR_summary_semanticpriming_lmerTest =   # Model with Kenward-Roger p values
  readRDS('semanticpriming/frequentist_analysis/results/KR_summary_semanticpriming_lmerTest.rds')
confint_semanticpriming_lmerTest =   # Confidence intervals
  readRDS('semanticpriming/frequentist_analysis/results/confint_semanticpriming_lmerTest.rds')

# Study 2: Semantic decision
# Code for data set below in 'semanticdecision/data/data_preparation' folder
semanticdecision = read.csv('semanticdecision/data/final_dataset/semanticdecision.csv')
# Code for models below in 'semanticdecision/frequentist_analysis' folder
semanticdecision_lmerTest =   # Primary model object
  readRDS('semanticdecision/frequentist_analysis/results/semanticdecision_lmerTest.rds')
KR_summary_semanticdecision_lmerTest =   # Model with Kenward-Roger p values
  readRDS('semanticdecision/frequentist_analysis/results/KR_summary_semanticdecision_lmerTest.rds')
confint_semanticdecision_lmerTest =   # Confidence intervals
  readRDS('semanticdecision/frequentist_analysis/results/confint_semanticdecision_lmerTest.rds')

# Study 3: Lexical decision
# Code for data set below in 'lexicaldecision/data/data_preparation' folder
lexicaldecision = read.csv('lexicaldecision/data/final_dataset/lexicaldecision.csv')
# Code for models below in 'lexicaldecision/frequentist_analysis' folder
lexicaldecision_lmerTest =   # Primary model object
  readRDS('lexicaldecision/frequentist_analysis/results/lexicaldecision_lmerTest.rds')
KR_summary_lexicaldecision_lmerTest =   # Model with Kenward-Roger p values
  readRDS('lexicaldecision/frequentist_analysis/results/KR_summary_lexicaldecision_lmerTest.rds')
confint_lexicaldecision_lmerTest =   # Confidence intervals
  readRDS('lexicaldecision/frequentist_analysis/results/confint_lexicaldecision_lmerTest.rds')

```

## Understanding the meaning of words

- Linguistic systems:

--

  - word co-occurrence

  - word association

--

- Embodied systems:

--

  - perceptual

  - motor

  - emotional

  - social

---

## Research question

How are the linguistic and the visual bases of conceptual processing manifested at the levels of participants, words and tasks?

--

- Three studies with large samples

  - semantic priming (Hutchison et al., 2013)

  - semantic decision (Pexman et al., 2017)

  - lexical decision (Balota et al., 2007)

--

- A-priori power analysis for each study

---

#### Main individual differences

  - vocabulary size

--

#### Main word-level variables

  - corpus-based linguistic measure in each study

  - visual strength (Lynott et al., 2020)

---

### Why vision?

Other studies have considered more perceptual modalities, such as hearing, touch, taste and smell (e.g., Bernabeu et al., 2017).

Vision most prominent in the brain (Reilly et al., 2020) and in language (Lynott et al., 2020; Winter et al., 2018).

Individual differences in vision easiest to measure (not present in these studies but planned for next study).

---

### Statistical analysis

#### Lexical covariates

- All possible lexical covariates out of 5 typical candidates:

  - Largest contributors without multicollinearity (*r* < .70) selected

#### Other important covariates for participant-level and word-level variables

- Word concreteness, fluid intelligence

---

## Semantic priming

### Primary data set: Hutchison et al. (2013)

```{r}

# Save number of participants and target_words
semanticpriming_number_participants = length(unique(semanticpriming$Participant))
semanticpriming_number_target_words = length(unique(semanticpriming$target_word))

# Number of target_words per participant
# Save mean as integer and SD rounded while keeping trailing zeros
semanticpriming_mean_target_words_per_participant = 
  semanticpriming %>% group_by(Participant) %>% 
  summarise(length(unique(target_word))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

semanticpriming_SD_target_words_per_participant = 
  semanticpriming %>% group_by(Participant) %>% 
  summarise(length(unique(target_word))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

# Number of participants per target_word
# Save mean as integer and SD rounded while keeping trailing zeros
semanticpriming_mean_participants_per_target_word = 
  semanticpriming %>% group_by(target_word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

semanticpriming_SD_participants_per_target_word = 
  semanticpriming %>% group_by(target_word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

```

- `r semanticpriming_number_participants` participants

- `r semanticpriming_number_target_words` target words

- `r nrow(semanticpriming) %>% formattable::comma(digits = 0)` observations

- `r semanticpriming_mean_target_words_per_participant` target words per participant (*SD* = `r semanticpriming_SD_target_words_per_participant`)

- `r semanticpriming_mean_participants_per_target_word` participants per target word (*SD* = `r semanticpriming_SD_participants_per_target_word`)

---

- Lexical decision = overt task

Over multiple trials, do the characters form a real word?

--

- Semantic priming = covert experimental design

>  <span style='color:blue;'> PRIME WORD  &nbsp; &nbsp;  target word</span>

>  &nbsp; &nbsp; &nbsp; &nbsp; PIKE  &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp;  prack

---

- Primes always real words; targets sometimes nonwords

  - Lexical decision on the target word (i.e., word or nonword?)

- Prime word: 150 ms

--

Interstimulus interval: 50 or 1,050 ms

Target word: until response or 3,000 ms

--

Linguistic information usually more influential than visual information

<div class='p-3 mb-2 bg-light text-dark' style='background-image: var(--bs-gradient);'>(Lam et al., 2015; Petilli et al., 2021)</div>

---

```{r rename-semanticpriming-effects, results = 'asis'}

# Rename effects using plain language and specify the random 
# slopes that were included. This is first done with the 
# main summary of the model (KR_summary...) and then with 
# the confidence intervals (confint...).

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_attentional_control'] =
  'Attentional control'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_vocabulary_size'] =
  'Vocabulary size <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_target_word_frequency'] =
  'Word frequency'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_target_number_syllables'] =
  'Number of syllables'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_word_concreteness_diff'] =
  'Word-concreteness difference'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_cosine_similarity'] =
  'Cosine similarity <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_visual_rating_diff'] =
  'Visual-strength difference <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_interstimulus_interval'] =
  'Interstimulus interval <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_word_concreteness_diff:z_vocabulary_size'] =
  'Word-concreteness difference : Vocabulary size'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_word_concreteness_diff:z_recoded_interstimulus_interval'] =
  'Word-concreteness difference : Interstimulus interval'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_word_concreteness_diff:z_recoded_participant_gender'] =
  'Word-concreteness difference : Gender'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_attentional_control:z_cosine_similarity'] =
  'Attentional control : Cosine similarity'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_attentional_control:z_visual_rating_diff'] =
  'Attentional control : Visual-strength difference'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_vocabulary_size:z_cosine_similarity'] =
  'Vocabulary size : Cosine similarity'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_vocabulary_size:z_visual_rating_diff'] =
  'Vocabulary size : Visual-strength difference'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_cosine_similarity'] =
  'Gender : Cosine similarity'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_visual_rating_diff'] =
  'Gender : Visual-strength difference'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_interstimulus_interval:z_cosine_similarity'] =
  'Interstimulus interval : Cosine similarity <sup>a</sup>'

rownames(KR_summary_semanticpriming_lmerTest$coefficients)[
  rownames(KR_summary_semanticpriming_lmerTest$coefficients) == 
    'z_recoded_interstimulus_interval:z_visual_rating_diff'] =
  'Interstimulus interval : Visual-strength difference <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_attentional_control'] =
  'Attentional control'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_vocabulary_size'] =
  'Vocabulary size <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_target_word_frequency'] =
  'Word frequency'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_target_number_syllables'] =
  'Number of syllables'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_word_concreteness_diff'] =
  'Word-concreteness difference'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_cosine_similarity'] =
  'Cosine similarity <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_visual_rating_diff'] =
  'Visual-strength difference <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_interstimulus_interval'] =
  'Interstimulus interval <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_word_concreteness_diff:z_vocabulary_size'] =
  'Word-concreteness difference : Vocabulary size'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_word_concreteness_diff:z_recoded_interstimulus_interval'] =
  'Word-concreteness difference : Interstimulus interval'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_word_concreteness_diff:z_recoded_participant_gender'] =
  'Word-concreteness difference : Gender'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_attentional_control:z_cosine_similarity'] =
  'Attentional control : Cosine similarity'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_attentional_control:z_visual_rating_diff'] =
  'Attentional control : Visual-strength difference'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_vocabulary_size:z_cosine_similarity'] =
  'Vocabulary size : Cosine similarity'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_vocabulary_size:z_visual_rating_diff'] =
  'Vocabulary size : Visual-strength difference'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_participant_gender:z_cosine_similarity'] =
  'Gender : Cosine similarity'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_participant_gender:z_visual_rating_diff'] =
  'Gender : Visual-strength difference'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_interstimulus_interval:z_cosine_similarity'] =
  'Interstimulus interval : Cosine similarity <sup>a</sup>'

rownames(confint_semanticpriming_lmerTest)[
  rownames(confint_semanticpriming_lmerTest) == 
    'z_recoded_interstimulus_interval:z_visual_rating_diff'] =
  'Interstimulus interval : Visual-strength difference <sup>a</sup>'

# Main effects (2/2)

# model_table() is a custom function from the 'R_functions' folder
semanticpriming_maineffects_2 =

  # model_table() is a custom function from the 'R_functions' folder
  model_table(KR_summary_semanticpriming_lmerTest,
              confidence_intervals = confint_semanticpriming_lmerTest,
              select_effects = c('Word-concreteness difference',
                                 'Cosine similarity <sup>a</sup>',
                                 'Visual-strength difference <sup>a</sup>',
                                 'Interstimulus interval <sup>a</sup>'),
              caption = '') %>%

  # Group predictors under headings
  pack_rows('Semantic variables', 1, 3) %>% 
  pack_rows('Task condition', 4, 4) %>%

  # Highlight covariates
  row_spec(1, background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', 
           general = "Shaded rows contain covariates. Variables involving 'difference' or 'similarity' are based on the comparison of the prime and the target words in each trial.",
           alphabet = c('By-participant random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

# Interactions (1/2)

# model_table() is a custom function from the 'R_functions' folder
semanticpriming_interactions_1 =

  # model_table() is a custom function from the 'R_functions' folder
  model_table(KR_summary_semanticpriming_lmerTest, 
              confidence_intervals = confint_semanticpriming_lmerTest,
              select_effects = c('Word-concreteness difference : Vocabulary size',
                                 'Word-concreteness difference : Interstimulus interval',
                                 'Word-concreteness difference : Gender',
                                 'Attentional control : Cosine similarity',
                                 'Attentional control : Visual-strength difference'),
              interaction_symbol_x = TRUE,
              caption = '') %>%

  # Highlight covariates
  row_spec(1:5, background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', 
           general = "All these interactions are covariates. Variables involving 'difference' or 'similarity' are based on the comparison of the prime and the target words in each trial.",
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

# Interactions (2/2)

# model_table() is a custom function from the 'R_functions' folder
semanticpriming_interactions_2 =

  # model_table() is a custom function from the 'R_functions' folder
  model_table(KR_summary_semanticpriming_lmerTest, 
              confidence_intervals = confint_semanticpriming_lmerTest,
              select_effects = c('Vocabulary size : Cosine similarity',
                                 'Vocabulary size : Visual-strength difference',
                                 'Gender : Cosine similarity',
                                 'Gender : Visual-strength difference',
                                 'Interstimulus interval : Cosine similarity <sup>a</sup>',
                                 'Interstimulus interval : Visual-strength difference <sup>a</sup>'),
              interaction_symbol_x = TRUE,
              caption = '') %>%

  footnote(general_title = 'Note. ', 
           general = "Variables involving 'difference' or 'similarity' are based on the comparison of the prime and the target words in each trial.",
           alphabet = c('By-participant random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

# Calculate R^2. This coefficient must be interpreted with caution 
# (Nakagawa et al., 2017; https://doi.org/10.1098/rsif.2017.0213). 
# Also, transform coefficient to rounded percentage.

Nakagawa2017_fixedeffects_R2_semanticpriming_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(semanticpriming_lmerTest)[1, 'R2m'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

Nakagawa2017_randomeffects_R2_semanticpriming_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(semanticpriming_lmerTest)[1, 'R2c'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

```

#### Predictors

.pull-left[

```{r semanticpriming-predictors-1, results = 'asis'}

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticpriming_lmerTest,
            confidence_intervals = confint_semanticpriming_lmerTest,
            select_effects = c('Attentional control',
                               'Vocabulary size <sup>a</sup>',
                               'Gender <sup>a</sup>',
                               'Word frequency',
                               'Number of syllables'),
            caption = '') %>%

  # Display the names of the effects only
  remove_column(2:6) %>%

  # Group predictors under headings
  pack_rows('Individual differences', 2, 4) %>% 
  pack_rows('Lexical covariates', 5, 6) %>%

  # Highlight covariates
  row_spec(c(2, 5:6), background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', general = "Shaded rows contain covariates.",
           alphabet = c('By-item random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

]

.pull-right[

```{r semanticpriming-predictors-2, results = 'asis'}

semanticpriming_maineffects_2 %>%
  remove_column(2:6)

```

]

---

<img src='img/Screenshot 2022-02-10 at 08.11.24.png' height='580px'/>

---

#### Results: Main effects (1/2)

```{r semanticpriming-maineffects-1, results = 'asis'}

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticpriming_lmerTest,
            confidence_intervals = confint_semanticpriming_lmerTest,
            select_effects = c('(Intercept)',
                               'Attentional control',
                               'Vocabulary size <sup>a</sup>',
                               'Gender <sup>a</sup>',
                               'Word frequency',
                               'Number of syllables'),
            caption = '') %>%

  # Group predictors under headings
  pack_rows('Individual differences', 2, 4) %>% 
  pack_rows('Lexical covariates', 5, 6) %>%

  # Highlight covariates
  row_spec(c(2, 5:6), background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', general = "Shaded rows contain covariates.",
           alphabet = c('By-item random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

---

#### Results: Main effects (2/2)

---

#### Results: Interactions (1/2)

---

#### Results: Interactions (2/2)

---

### Variance explained (Nakagawa et al., 2017)

- the fixed effects explained `r Nakagawa2017_fixedeffects_R2_semanticpriming_lmerTest`

- the random effects explained `r Nakagawa2017_randomeffects_R2_semanticpriming_lmerTest`

---

### Power analysis

---

## Semantic decision

### Primary data set: Pexman et al. (2017)

```{r}

# Save number of participants and words
semanticdecision_number_participants = length(unique(semanticdecision$Participant))
semanticdecision_number_words = length(unique(semanticdecision$Word))

# Number of words per participant
# Save mean as integer and SD rounded while keeping trailing zeros
semanticdecision_mean_words_per_participant = 
  semanticdecision %>% group_by(Participant) %>% 
  summarise(length(unique(Word))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

semanticdecision_SD_words_per_participant = 
  semanticdecision %>% group_by(Participant) %>% 
  summarise(length(unique(Word))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

# Number of participants per word
# Save mean as integer and SD rounded while keeping trailing zeros
semanticdecision_mean_participants_per_word = 
  semanticdecision %>% group_by(Word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

semanticdecision_SD_participants_per_word = 
  semanticdecision %>% group_by(Word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

```

- `r semanticdecision_number_participants` participants

- `r semanticdecision_number_words` words

- `r nrow(semanticdecision) %>% formattable::comma(digits = 0)` observations

- `r semanticdecision_mean_words_per_participant` words per participant (*SD* = `r semanticdecision_SD_words_per_participant`)

- `r semanticdecision_mean_participants_per_word` participants per word (*SD* = `r semanticdecision_SD_participants_per_word`)

---

#### Predictors

<img src='img/Screenshot 2022-02-10 at 08.11.48.png' height='300px'/>

---

<img src='img/Screenshot 2022-02-10 at 08.12.05.png' height='430px'/>

---

#### Results: Main effects (1/2)

```{r semanticdecision-maineffects-1, results = 'asis'}

# Rename effects using plain language and specify the random 
# slopes that were included.

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_information_uptake'] = 
  'Information uptake'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_vocabulary_size'] = 
  'Vocabulary size <sup>a</sup>'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_information_uptake'] = 
  'Information uptake'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_vocabulary_size'] = 
  'Vocabulary size <sup>a</sup>'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticdecision_lmerTest, 
            confidence_intervals = confint_semanticdecision_lmerTest,
            select_effects = c('(Intercept)',
                               'Information uptake',
                               'Vocabulary size <sup>a</sup>',
                               'Gender <sup>a</sup>'),
            caption = '') %>%

  # Group predictors under headings
  pack_rows('Individual differences', 2, 4) %>%

  # Highlight covariates
  row_spec(2, background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', general = 'Shaded rows contain covariates.',
           alphabet = c('By-item random slopes were included for this effect. '),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

---

#### Results: Main effects (2/2)

```{r semanticdecision-maineffects-2, results = 'asis'}

# Rename effects using plain language and specify the random 
# slopes that were included.

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_word_frequency'] = 
  'Word frequency'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_orthographic_Levenshtein_distance'] = 
  'OLD20'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_word_concreteness'] = 
  'Word concreteness'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_word_cooccurrence'] = 
  'Word co-occurrence <sup>a</sup>'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_visual_rating'] =
  'Visual strength <sup>a</sup>'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_word_frequency'] = 
  'Word frequency'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_orthographic_Levenshtein_distance'] = 
  'OLD20'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_word_concreteness'] = 
  'Word concreteness'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_word_cooccurrence'] = 
  'Word co-occurrence <sup>a</sup>'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_visual_rating'] =
  'Visual strength <sup>a</sup>'

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticdecision_lmerTest, 
            confidence_intervals = confint_semanticdecision_lmerTest,
            select_effects = c('Word frequency',
                               'OLD20',
                               'Word concreteness',
                               'Word co-occurrence <sup>a</sup>',
                               'Visual strength <sup>a</sup>'),
            caption = '') %>%

  # Group predictors under headings
  pack_rows('Lexical covariates', 1, 2) %>% 
  pack_rows('Semantic variables', 3, 5) %>%

  # Highlight covariates
  row_spec(1:3, background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', 
           general = 'Shaded rows contain covariates. OLD = Orthographic Levenshtein distance.',
           alphabet = c('By-participant random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

---

#### Results: Interactions (1/2)

```{r semanticdecision-interactions-1, results = 'asis'}

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_word_concreteness:z_vocabulary_size'] =
  'Word concreteness : Vocabulary size'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_word_concreteness:z_recoded_participant_gender'] =
  'Word concreteness : Gender'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_information_uptake:z_word_cooccurrence'] =
  'Information uptake : Word co-occurrence'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_information_uptake:z_visual_rating'] =
  'Information uptake : Visual strength'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_word_concreteness:z_vocabulary_size'] =
  'Word concreteness : Vocabulary size'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_word_concreteness:z_recoded_participant_gender'] =
  'Word concreteness : Gender'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_information_uptake:z_word_cooccurrence'] =
  'Information uptake : Word co-occurrence'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_information_uptake:z_visual_rating'] =
  'Information uptake : Visual strength'

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticdecision_lmerTest, 
            confidence_intervals = confint_semanticdecision_lmerTest,
            select_effects = c('Word concreteness : Vocabulary size',
                               'Word concreteness : Gender',
                               'Information uptake : Word co-occurrence',
                               'Information uptake : Visual strength'),
            interaction_symbol_x = TRUE,
            caption = '') %>%

  footnote(general_title = 'Note. ', 
           general = 'All these interactions are covariates.',
           footnote_as_chunk = TRUE) %>%

  # Highlight covariates
  row_spec(1:4, background = '#EFEFEF') %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

---

#### Results: Interactions (2/2)

```{r semanticdecision-interactions-2, results = 'asis'}

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_vocabulary_size:z_word_cooccurrence'] =
  'Vocabulary size : Word co-occurrence'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_vocabulary_size:z_visual_rating'] =
  'Vocabulary size : Visual strength'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_word_cooccurrence'] =
  'Gender : Word co-occurrence'

rownames(KR_summary_semanticdecision_lmerTest$coefficients)[
  rownames(KR_summary_semanticdecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_visual_rating'] =
  'Gender : Visual strength'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_vocabulary_size:z_word_cooccurrence'] =
  'Vocabulary size : Word co-occurrence'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_vocabulary_size:z_visual_rating'] =
  'Vocabulary size : Visual strength'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_recoded_participant_gender:z_word_cooccurrence'] =
  'Gender : Word co-occurrence'

rownames(confint_semanticdecision_lmerTest)[
  rownames(confint_semanticdecision_lmerTest) == 
    'z_recoded_participant_gender:z_visual_rating'] =
  'Gender : Visual strength'

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_semanticdecision_lmerTest, 
            confidence_intervals = confint_semanticdecision_lmerTest,
            select_effects = c('Vocabulary size : Word co-occurrence',
                               'Vocabulary size : Visual strength',
                               'Gender : Word co-occurrence',
                               'Gender : Visual strength'),
            interaction_symbol_x = TRUE, 
            caption = '') %>%

  kable_classic(full_width = F, html_font = 'Cambria')

# Calculate R^2. This coefficient must be interpreted with caution 
# (Nakagawa et al., 2017; https://doi.org/10.1098/rsif.2017.0213). 
# Also, transform coefficient to rounded percentage.

Nakagawa2017_fixedeffects_R2_semanticdecision_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(semanticdecision_lmerTest)[1, 'R2m'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

Nakagawa2017_randomeffects_R2_semanticdecision_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(semanticdecision_lmerTest)[1, 'R2c'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

```

---

### Variance explained (Nakagawa et al., 2017)

- the fixed effects explained `r Nakagawa2017_fixedeffects_R2_semanticdecision_lmerTest`

- the random effects explained `r Nakagawa2017_randomeffects_R2_semanticdecision_lmerTest`

---

### Interactions

```{r}

# Run plot through source() rather than directly in this R Markdown document 
# to preserve the italics in the legend.

source('semanticpriming/frequentist_analysis/semanticpriming-interaction-vocabulary-cosinesimilarity.R', 
       local = TRUE)

include_graphics(
  paste0(
    getwd(),  # Keeping illegal characters out of include_graphics()
    '/semanticpriming/frequentist_analysis/plots/semanticpriming-interaction-vocabulary-cosinesimilarity.pdf'
  ))

```

---

### Power analysis

---

## Lexical decision

### Primary data set: Balota et al. (2007)

```{r}

# Save number of participants and words
lexicaldecision_number_participants = length(unique(lexicaldecision$Participant))
lexicaldecision_number_words = length(unique(lexicaldecision$word))

# Number of words per participant
# Save mean as integer and SD rounded while keeping trailing zeros
lexicaldecision_mean_words_per_participant = 
  lexicaldecision %>% group_by(Participant) %>% 
  summarise(length(unique(word))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

lexicaldecision_SD_words_per_participant = 
  lexicaldecision %>% group_by(Participant) %>% 
  summarise(length(unique(word))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

# Number of participants per word
# Save mean as integer and SD rounded while keeping trailing zeros
lexicaldecision_mean_participants_per_word = 
  lexicaldecision %>% group_by(word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% mean %>% round(0)

lexicaldecision_SD_participants_per_word = 
  lexicaldecision %>% group_by(word) %>% 
  summarise(length(unique(Participant))) %>% 
  select(2) %>% unlist %>% sd %>% sprintf('%.2f', .)

```

- `r lexicaldecision_number_participants` participants

- `r lexicaldecision_number_words` words

- `r nrow(lexicaldecision) %>% formattable::comma(digits = 0)` observations

- `r lexicaldecision_mean_words_per_participant` words per participant (*SD* = `r lexicaldecision_SD_words_per_participant`)

- `r lexicaldecision_mean_participants_per_word` participants per word (*SD* = `r lexicaldecision_SD_participants_per_word`)

---

#### Predictors

<img src='img/Screenshot 2022-02-10 at 08.12.25.png' height='270px'/>

---

<img src='img/Screenshot 2022-02-10 at 08.12.37.png' height='400px'/>

---

#### Results: Main effects

```{r lexicaldecision-maineffects, results = 'asis'}

# Rename effects using plain language and specify the random 
# slopes that were included.

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_vocabulary_age'] =
  'Vocabulary age <sup>a</sup>'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_orthographic_Levenshtein_distance'] =
  'OLD20'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_word_concreteness'] =
  'Word concreteness'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_word_frequency'] =
  'Word frequency <sup>b</sup>'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_visual_rating'] =
  'Visual strength <sup>b</sup>'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_vocabulary_age'] =
  'Vocabulary age <sup>a</sup>'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_recoded_participant_gender'] =
  'Gender <sup>a</sup>'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_orthographic_Levenshtein_distance'] =
  'OLD20'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_word_concreteness'] =
  'Word concreteness'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_word_frequency'] =
  'Word frequency <sup>b</sup>'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_visual_rating'] =
  'Visual strength <sup>b</sup>'

# model_table() is a custom function from the 'R_functions' folder
model_table(KR_summary_lexicaldecision_lmerTest, 
            confidence_intervals = confint_lexicaldecision_lmerTest,
            select_effects = c('(Intercept)',
                               'Vocabulary age <sup>a</sup>',
                               'Gender <sup>a</sup>',
                               'OLD20',
                               'Word concreteness',
                               'Word frequency <sup>b</sup>',
                               'Visual strength <sup>b</sup>'),
            caption = '') %>%

  # Group predictors under headings
  pack_rows('Individual differences', 2, 3) %>% 
  pack_rows('Lexical-semantic covariates', 4, 5) %>% 
  pack_rows('Lexical-semantic variables', 6, 7) %>%

  # Highlight covariates
  row_spec(4:5, background = '#EFEFEF') %>%

  footnote(general_title = 'Note. ', 
           general = 'Shaded rows contain covariates. OLD = Orthographic Levenshtein distance.',
           alphabet = c('By-item random slopes were included for this effect. ', 
                        'By-participant random slopes were included for this effect.'),
           footnote_as_chunk = TRUE) %>%

  kable_classic(full_width = F, html_font = 'Cambria')

```

---

#### Results: Interactions

```{r lexicaldecision-interactions, results = 'asis'}

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_word_concreteness:z_vocabulary_age'] =
  'Word concreteness : Vocabulary age'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_word_concreteness:z_recoded_participant_gender'] =
  'Word concreteness : Gender'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_vocabulary_age:z_word_frequency'] =
  'Vocabulary age : Word frequency'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_vocabulary_age:z_visual_rating'] =
  'Vocabulary age : Visual strength'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_word_frequency'] =
  'Gender : Word frequency'

rownames(KR_summary_lexicaldecision_lmerTest$coefficients)[
  rownames(KR_summary_lexicaldecision_lmerTest$coefficients) == 
    'z_recoded_participant_gender:z_visual_rating'] =
  'Gender : Visual strength'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_word_concreteness:z_vocabulary_age'] =
  'Word concreteness : Vocabulary age'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_word_concreteness:z_recoded_participant_gender'] =
  'Word concreteness : Gender'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_vocabulary_age:z_word_frequency'] =
  'Vocabulary age : Word frequency'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_vocabulary_age:z_visual_rating'] =
  'Vocabulary age : Visual strength'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_recoded_participant_gender:z_word_frequency'] =
  'Gender : Word frequency'

rownames(confint_lexicaldecision_lmerTest)[
  rownames(confint_lexicaldecision_lmerTest) == 
    'z_recoded_participant_gender:z_visual_rating'] =
  'Gender : Visual strength'

model_table(KR_summary_lexicaldecision_lmerTest, 
            confidence_intervals = confint_lexicaldecision_lmerTest,
            select_effects = c('Word concreteness : Vocabulary age',
                              'Word concreteness : Gender',
                              'Vocabulary age : Word frequency',
                              'Vocabulary age : Visual strength',
                              'Gender : Word frequency',
                              'Gender : Visual strength'),
            interaction_symbol_x = TRUE,
            caption = '') %>%

  footnote(general_title = 'Note. ', 
           general = 'Shaded rows contain covariates.',
           footnote_as_chunk = TRUE) %>%

  # Highlight covariates
  row_spec(1:2, background = '#EFEFEF') %>%

  kable_classic(full_width = F, html_font = 'Cambria')

# Calculate R^2. This coefficient must be interpreted with caution 
# (Nakagawa et al., 2017; https://doi.org/10.1098/rsif.2017.0213). 
# Also, transform coefficient to rounded percentage.

Nakagawa2017_fixedeffects_R2_lexicaldecision_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(lexicaldecision_lmerTest)[1, 'R2m'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

Nakagawa2017_randomeffects_R2_lexicaldecision_lmerTest = 
  paste0(
    (MuMIn::r.squaredGLMM(lexicaldecision_lmerTest)[1, 'R2c'][[1]] * 100) %>% 
      sprintf('%.2f', .),
    '%')

```

---

### Variance explained (Nakagawa et al., 2017)

- the fixed effects explained `r Nakagawa2017_fixedeffects_R2_lexicaldecision_lmerTest`

- the random effects explained `r Nakagawa2017_randomeffects_R2_lexicaldecision_lmerTest`

---

### Power analysis

---

# References

Albers, C., & Lakens, D. (2018). When power analyses based on pilot data are biased: Inaccurate effect size estimators and follow-up bias. *Journal of Experimental Social Psychology, 74*, 187–195. https://doi.org/10.1016/j.jesp.2017.09.004

Balota, D. A., Yap, M. J., Hutchison, K. A., Cortese, M. J., Kessler, B., Loftis, B., Neely, J. H., Nelson, D. L., Simpson, G. B., & Treiman, R. (2007). The English Lexicon Project. *Behavior Research Methods, 39*, 445–459. https://doi.org/10.3758/BF03193014

Bernabeu, P., Lynott, D., & Connell, L. (2021). Preregistration: *The interplay between linguistic and embodied systems in conceptual processing*. OSF. https://osf.io/ftydw/

Bernabeu, P., Willems, R. M., & Louwerse, M. M. (2017). Modality switch effects emerge early and increase throughout conceptual processing: Evidence from ERPs. In G. Gunzelmann, A. Howes, T. Tenbrink, & E. J. Davelaar (Eds.), *Proceedings of the 39th Annual Conference of the Cognitive Science Society* (pp. 1629-1634). Austin, TX: Cognitive Science Society. https://mindmodeling.org/cogsci2017/papers/0318

---

Brauer, M., & Curtin, J. J. (2018). Linear mixed-effects models and the analysis of nonindependent data: A unified framework to analyze categorical and continuous independent variables that vary within-subjects and/or within-items. *Psychological Methods, 23*(3), 389–411. https://doi.org/10.1037/met0000159

Hutchison, K. A., Balota, D. A., Neely, J. H., Cortese, M. J., Cohen-Shikora, E. R., Tse, C.-S., Yap, M. J., Bengson, J. J., Niemeyer, D., & Buchanan, E. (2013). The semantic priming project. *Behavior Research Methods, 45*, 1099–1114. https://doi.org/10.3758/s13428-012-0304-z

Lam, K. J., Dijkstra, T., & Rueschemeyer, S. A. (2015). Feature activation during word recognition: Action, visual, and associative-semantic priming effects. *Frontiers in Psychology, 6*, 659. https://doi.org/10.3389/fpsyg.2015.00659

Loken, E., & Gelman, A. (2017). Measurement error and the replication crisis. *Science, 355*(6325), 584-585. https://doi.org/10.1126/science.aal3618

Luke, S. G. (2017). Evaluating significance in linear mixed-effects models in R. *Behavior Research Methods, 49*(4), 1494–1502. https://doi.org/10.3758/s13428-016-0809-y

---

Lynott, D., Connell, L., Brysbaert, M., Brand, J., & Carney, J. (2020). The Lancaster Sensorimotor Norms: Multidimensional measures of perceptual and action strength for 40,000 English words. *Behavior Research Methods, 52*, 1271–1291. https://doi.org/10.3758/s13428-019-01316-z.

Petilli, M. A., Günther, F., Vergallito, A., Ciapparelli, M., & Marelli, M. (2021). Data-driven computational models reveal perceptual simulation in word processing. *Journal of Memory and Language, 117*, 104194. https://doi.org/10.1016/j.jml.2020.104194

Pexman, P. M., Heard, A., Lloyd, E., & Yap, M. J. (2017). The Calgary semantic decision project: Concrete/abstract decision data for 10,000 English words. *Behavior Research Methods, 49*(2), 407–417. https://doi.org/10.3758/s13428-016-0720-6

Pexman, P. M., & Yap, M. J. (2018). Individual differences in semantic processing: Insights from the Calgary semantic decision project. *Journal of Experimental Psychology: Learning, Memory, and Cognition, 44*(7), 1091–1112. https://doi.org/10.1037/xlm0000499

---

Reilly, J., Flurie, M., & Peelle, J. E. (2020). The English lexicon mirrors functional brain
activation for a sensory hierarchy dominated by vision and audition: Point-counterpoint. *Journal of Neurolinguistics, 55*, 100895. https://doi.org/10.1016/j.jneuroling.2020.100895

Singmann, H., & Kellen, D. (2019). *An Introduction to Mixed Models for Experimental Psychology*. In D. H. Spieler & E. Schumacher (Eds.), New Methods in Cognitive Psychology (pp. 4--31). Hove, UK: Psychology Press. http://singmann.org/download/publications/singmann_kellen-introduction-mixed-models.pdf

Tulving, E. (2007). Are There 256 Different Kinds of Memory? In J. S. Nairne (Ed.), *The foundations of remembering: Essays in honor of Henry L. Roediger, III* (p. 39–52). Psychology Press.

Wingfield, C., & Connell, L. (2019). *Understanding the role of linguistic distributional knowledge in cognition.* PsyArXiv. https://doi.org/10.31234/osf.io/hpm4z

---

Yap, M. J., Balota, D. A., Sibley, D. E., & Ratcliff, R. (2012). Individual differences in visual word recognition: Insights from the English Lexicon Project. *Journal of Experimental Psychology: Human Perception and Performance*, 38, 1, 53–79. https://doi.org/10.1037/a0024177

Yap, M. J., Hutchison, K. A., & Tan, L. C. (2017). Individual differences in semantic priming performance: Insights from the semantic priming project. In M. N. Jones (Ed.), *Frontiers of cognitive psychology. Big data in cognitive science* (p. 203–226). Routledge/Taylor & Francis Group.

Yap, M. J., Tse, C.-S., & Balota, D. A. (2009). Individual differences in the joint effects of semantic priming and word frequency revealed by RT distributional analyses: The role of lexical integrity. *Journal of Memory and Language, 61*(3), 303–325. https://doi.org/10.1016/j.jml.2009.07.001

Winter, B., Perlman, M., & Majid, A. (2018). Vision dominates in perceptual language: English sensory vocabulary is optimized for usage. *Cognition, 179*, 213–220. https://doi.org/10.1016/j.cognition.2018.05.008